---
title: "Using the iPCQ calculation function"
author: "Stijn Peeters"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the iPCQ calculation function}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# Purpose

`calculate_ipcq()` calculates productivity losses from iPCQ-derived data. It
returns separate hours and costs for short absenteeism, long absenteeism,
presenteeism, and unpaid work. It never requires identifying information.

# Required input

Each row represents one participant at one measurement. Run:

```{r, eval=FALSE}
library(TforT)
ipcq_input_spec()
```

The required columns are:

| Column | Meaning |
|---|---|
| `date_of_this_measurement` | Completion date in `YYYY-MM-DD` format |
| `hours_work_week` | Contracted paid-work hours per week |
| `days_work_week` | Paid working days per week |
| `days_sick` | Working days absent during the input period |
| `sick_longer_than_4_weeks` | `1` for uninterrupted long absence; otherwise `0` |
| `date_start_sickness` | Start date; required for long absence |
| `days_suffering_from_problems` | Days worked with health problems |
| `rate_of_work` | Productivity from 0 (none) to 10 (normal) |
| `days_less_unpaid_work` | Days with reduced unpaid work |
| `average_hours_unpaid_work` | Average unpaid-work hours lost on those days |

Extra columns, such as participant ID, arm, or time point, are retained.

# Core calculations

For short absence:

```
hours per working day = hours_work_week / days_work_week
short absence hours   = days_sick * hours per day * projection factor
short absence costs   = short absence hours * paid-work hourly price
```

For presenteeism:

```
presenteeism hours = affected working days * hours per day *
                     (1 - rate_of_work / 10) * projection factor
```

For unpaid work:

```
unpaid-work hours = affected days * average lost hours * projection factor
```

The projection factor is `target_period_weeks / input_period_weeks`. Long
absence is not projected from `days_sick`: its valued duration is the overlap
between the target period, sickness spell, and friction window.

# Dutch reference prices

Euro prices:

```{r, eval=FALSE}
result <- calculate_ipcq(data, reference_year = 2024, currency = "EUR")
```

International dollars:

```{r, eval=FALSE}
result <- calculate_ipcq(data, reference_year = 2024, currency = "INT$")
```

In both cases the paid-work price, unpaid-work price, and Dutch friction period
come from `tatooheene`.

# Custom country assumptions

```{r, eval=FALSE}
result <- calculate_ipcq(
  data,
  reference_year = 2024,
  currency = "GBP",
  paid_work_hour_price = 20,
  unpaid_work_hour_price = 8,
  friction_period_days = 70
)
```

For another country, provide both hourly prices and an appropriate friction
period. Record the source and price year in the study documentation.

# Output

The most important added columns are:

```text
absenteeism_short_hours       absenteeism_short_costs
absenteeism_long_hours        absenteeism_long_costs
absenteeism_hours             absenteeism_costs
presenteeism_hours            presenteeism_costs
unpaid_work_hours             unpaid_work_costs
total_productivity_costs
```

The output also records the prices, currency, friction period, extrapolation
factor, and whether assumptions came from `tatooheene` or the user.

# Manual verification and tests

Locate the bundled manual cases:

```{r, eval=FALSE}
system.file("extdata", "ipcq_manual_test_cases.csv", package = "TforT")
```

During package development, run:

```{r, eval=FALSE}
devtools::load_all()
devtools::test()
devtools::check()
```

# Missing values and validation

The default `na_action = "error"` is recommended. It prevents an incomplete
answer from being silently treated as no productivity loss. A sickness start
date may be missing only for respondents without long absence.
