---
title: "Using the iPCQ Shiny calculator"
author: "Stijn Peeters"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the iPCQ Shiny calculator}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# Start the application

Install and load TforT, then run:

```{r, eval=FALSE}
library(TforT)
run_ipcq_app()
```

# Prepare a file

The app accepts CSV, XLSX, and RDS data frames. Click **Download input
template** in the app for the correct headers. Each row is one participant and
measurement. Extra columns are retained.

Dates should use `YYYY-MM-DD`. Leave `date_start_sickness` empty only when
`sick_longer_than_4_weeks` is zero. Do not upload names, addresses, contact
information, or other directly identifying information.

# Choose calculation settings

1. Set the number of weeks represented by the answers. Standard iPCQ responses
   normally represent four weeks.
2. Set the target period. A longer target period projects short absence,
   presenteeism, and unpaid work using the ratio of target to input weeks.
3. Select a reference year.
4. Choose Dutch reference prices or custom prices.
5. Choose the Dutch friction period or provide a custom number of calendar days.
6. Click **Calculate costs**.

# Price options

With **Dutch reference prices**, select either EUR or INT$. The app obtains paid
and unpaid hourly prices from `tatooheene`.

With **Provide my own prices**, enter a currency label and both hourly prices.
For non-Dutch settings, normally also select **Provide my own period** and enter
the relevant national friction period.

# Review and download

The **Results** tab contains original variables and calculated components. The
**Calculation settings** tab records prices, currency, friction period, sources,
and the projection factor. Export the result as CSV, Excel, or RDS.

# Interpreting errors

The app reports missing columns, invalid dates, negative values, impossible
numbers of working days, missing long-absence dates, and unsupported price
combinations. Correct the source file and upload it again; calculations never
alter the original uploaded file.

# Reproducibility

Archive the result together with the original input, package version,
questionnaire version, selected price year, currency, and source of custom
assumptions. Projection beyond the questionnaire's observed period should be
reported as an analytical assumption.
